---
title: "Assignment4"
author: "Swetha"
date: "3/14/2022"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
a.Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on.
```{r}
#Loading the Required packages
library(flexclust)
library(cluster)
library(tidyverse)  
library(factoextra) 
library(FactoMineR)
library(ggcorrplot)
#loading the data
getwd()
setwd("C:/Users/mercy/OneDrive/Desktop/FML/Assignment4")
Info<- read.csv("Pharmaceuticals.csv")
# I am selecting columns from 3 to 11 and storing the data in variable Info1
Info1 <- Info[3:11]
# Using head function to display the first 6 rows of data
head(Info1)
summary(Info1)
```
# The variables are measured in different weights throughout the rows, we will scale the data in Info1 and save the scaled data in the Info2 dataframe. I am Calculating the distance between the rows of data and visualizing the distance matrix using get_dist and fviz_dist functions which are available in factoextra package.
```{r}
Info2 <- scale(Info1)
row.names(Info2) <- Info[,1]
distance <- get_dist(Info2)
fviz_dist(distance)
```
# Now I am creating the correlation Matrix and I am printing to check the correlation among major variables
```{r}
corr <- cor(Info2)
ggcorrplot(corr, outline.color = "grey50", lab = TRUE, hc.order = TRUE, type = "full") 
```
#The Correlation Matrix reveals that ROA, ROE, Net Profit Margin, and Market Cap is high. By using  Principal Component Analysis I am finding out weightage of major variables in the data set. Here I am assuming best number of cluster is 5.
```{r}
pca <- PCA(Info2)
var <- get_pca_var(pca)
fviz_pca_var(pca, col.var="contrib",
             gradient.cols = c("grey","yellow","purple","red","blue"),ggrepel = TRUE ) + labs( title = "PCA Variable Variance")
```
We may deduce from PCA Variable Variance that ROA,ROE, Net Profit Margin, Market Cap, and Asset Turnover contribute over 61% to the two PCA components/dimensions (Variables)and I am using elbow method to find optimal number of customers.
```{r}
set.seed(10)
wss <- vector()
for(i in 1:10) wss[i] <- sum(kmeans(Info2,i)$withinss)
plot(1:10, wss , type = "b" , main = paste('Cluster of Companies') , xlab = "Number of Clusters", ylab="wss")
wss
```
I got the same number as assumed. Optimal cluster is at 5 .
## Silhouette Method
Finding best number of clusters.
```{r}
fviz_nbclust(Info2, kmeans, method = "silhouette")
```
Here also the idealnumber of clusters is 5. Using k-means algorithm to cluster with 5.
```{r}
set.seed(1)
k5 <- kmeans(Info2, centers = 5, nstart = 25) # k = 5, number of restarts = 25
k5$centers 
k5$size  
fviz_cluster(k5, data = Info2) 
```
#kmeans clustering, using Manhattan Distance
```{r}
set.seed(1)
k51 = kcca(Info2, k=5, kccaFamily("kmedians"))
k51
#Using predict function.
clusters_index <- predict(k51)
dist(k51@centers)
image(k51)
points(Info2, col=clusters_index, pch=19, cex=0.3)
```
b.Interpret the clusters with respect to the numerical variables used in forming the clusters
Using Kmeans method to calculate Mean.
```{r}
Info1 %>% mutate(Cluster = k5$cluster) %>% group_by(Cluster) %>% summarise_all("mean")
clusplot(Info2,k5$cluster, main="Clusters",color = TRUE, labels = 2,lines = 0)
```
Comapnies are categorized into different clusters as follows:

Cluster 1: ELN, MRX, WPI and AVE
Cluster 2: AGN and PHA
Cluster 3: AHM,WYE,BMY,AZN, LLY, ABT, NVS and SGP
Cluster 4: BAY, CHTT and IVX
Cluster 5: JNJ, MRK, PFE and GSK
From the means of the cluster variables , we can say that,
Cluster 1 has the fastest revenue growth, the highest Net Profit Margin, and the lowest PE ratio. It can be purchased or held in reserve..Cluster 2 PE ratio is very high Cluster 3 has average risk Cluster 4 Though it has a good PE ratio, it carries a very high risk , very very high leverage and low Net Profit margin , making it very risky to own. Revenue growth is also very low.Cluster 5 has a high market capitalization, return on investment, return on assets, asset turnover, and net profit margin. With a low PE ratio, the stock price is moderately valued and hence can be purchased and held evenue growth of 18.5% is good.
c.Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used informing the clusters)
#plotting clusters against the variables to check for any patterns
```{r}
Info3 <- Info[12:14] %>% mutate(Clusters=k5$cluster)
ggplot(Info3, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters')
ggplot(Info3, mapping = aes(factor(Clusters),fill = Location))+geom_bar(position = 'dodge')+labs(x ='Clusters')
ggplot(Info3, mapping = aes(factor(Clusters),fill = Exchange))+geom_bar(position = 'dodge')+labs(x ='Clusters')
```

-->.Clusters and the variable Median Recommendation appear to follow a pattern.
-->Except for the fact that the bulk of the clusters/companies are listed on the NYSE and are based in the United States, there appears to be no discernible pattern among the clusters, locations, or exchanges.
d.Provide an appropriate name for each cluster using any or all of the variables in the dataset.
Cluster 1: Best Buying
Cluster 2: Highly Risky
Cluster 3: Go for it 
Cluster 4: Very Risky or Runaway 
Cluster 5: Ideal to Own 


